Paper 2 !DRAFT! Do not circulate!

### 2014-11-07 Articles/Article_EcolEcon_2014/Paper2.Rnw
Now making lambda time series graphs.

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconModels)
require(EconData)
require(ggplot2)
require(plyr)
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
require(knitr)

@@

<<data, include=FALSE, eval=TRUE>>=
# Load historical data
historicalData <- rbind.fill(REXS1960, IST, Leeds) # From the EconData package

@@

@@

>>>>

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconModels)
require(EconData)
require(ggplot2)
require(plyr)
require(dplyr) # Load in this order or the mutate and arrange functions will not work as desired.
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
require(knitr)
require(mosaic)

@@

<<data, include=FALSE>>=
# Load historical data
historicalData <- rbind.fill(REXS1960, IST, Leeds) # From the EconData package

@@

The term $\mathrm{e}^{\lambda (t-t_0)}$ is known as the Solow Residual ($A$). 
Assuming that parameters $\theta$, $\alpha$, and $\beta$ are known,
we can estimate the value of $\lambda$ as follows.
First, assume constant returns to scale such that $\gamma = 1 - \alpha - \beta$.
Then, take the natural logarithm ($\ln$) of Equation~\ref{eq:CD} to obtain
%
\begin{equation} \label{eq:lnCD}
  \ln y = \ln \theta + \ln A + \alpha \ln k + \beta \ln l + (1 - \alpha - \beta) \ln e \;.
\end{equation}
%
By taking the derivative of Equation~\ref{eq:lnCD} with respect to time ($t$) 
and noting that model parameters $\theta$ and $\alpha$ are 
constant with respect to time, we obtain
%
\begin{equation} \label{eq:ddtlnCD}
  \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
      = \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
      + \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
      + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
      + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} \;.
\end{equation}
%
Solving for the Solow Residual term gives
%
\begin{equation} \label{eq:solow_resid}
  \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
      = \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
      - \left[
              \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
              + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
              + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} 
        \right] \;.
\end{equation}
%
Recognizing that $A \equiv \mathrm{e}^{\lambda (t-t_0)}$ gives
%
\begin{equation}
  \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
    = \frac{1}{\mathrm{e}^{\lambda (t-t_0)}} \; \lambda \mathrm{e}^{\lambda (t-t_0)}
    = \lambda \; ,
\end{equation}
%
which can be substituted into Equation~\ref{eq:solow_resid} to find
%
\begin{equation} \label{eq:lambda}
  \lambda 
    = \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
    - \left[
              \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
              + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
              + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} 
      \right] \;.
\end{equation}
%
Equation~\ref{eq:lambda} applies for any instant in time.

If we approximate derivatives in Equation~\ref{eq:lambda} with forward differences
between times $i$ and $j$, we find
%
\begin{equation} \label{eq:lambdaij}
  \lambda_{i,j}
    \approx \frac{1}{y_i}\frac{y_j - y_i}{t_j - t_i} 
    - \left[
            \alpha \frac{1}{k_i}\frac{k_j - k_i}{t_j - t_i} 
            + \beta \frac{1}{l_i}\frac{l_j - l_i}{t_j - t_i} 
            + (1 - \alpha - \beta) \frac{1}{e_i}\frac{e_j - e_i}{t_j - t_i} 
      \right] \;,
\end{equation}
% 
where $\lambda_{i,j}$ approximates the true, instantaneous value of $\lambda$
between times $i$ and $j$.

Interestingly, 
%
\begin{equation}
  \lambda_{1,n}
    \neq \sum_{i=2}^{n} \lambda_{i,i-1}  \;.
\end{equation}
% 
Rather,
%
\begin{equation} \label{eq:lambda1n}
  \lambda_{1,n}
    \approx \frac{1}{y_1}\frac{y_n - y_1}{t_n - t_1} 
      - \left[ 
            \alpha \frac{1}{k_1}\frac{k_n - k_1}{t_n - t_1} 
            + \beta \frac{1}{l_1}\frac{l_n - l_1}{t_n - t_1} 
            + (1 - \alpha - \beta) \frac{1}{e_1}\frac{e_n - e_1}{t_n - t_1} 
      \right] \; ,
\end{equation}
% 
which will become increasingly inaccurate over large time spans,
because $y_1$, $k_1$, $l_1$, and $e_1$ will be less representative
of the average value of $y$, $k$, $l$, and $e$ in the time span, respectively.
This suggests that an averaging approach such as
%
\begin{equation}
  \lambda_{1,n}
    \approx \frac{\sum_{i=2}^{n} \lambda_{i,i-1}}{n-1}
\end{equation}
%
is a better approximation of Equation~\ref{eq:lambda1n},
which is, itself, an approximation of the true value of $\lambda$ 
given by Equation~\ref{eq:lambda}.

@@

Given the estimates for $\alpha$, $\beta$, and $\gamma$ for the Cobb-Douglas model,
we can estimate values for $\lambda$ on a year-by-year basis using Equation~\ref{eq:lambdaij}.
The results are shown in the Figure~\ref{fig:CD_lambda_time_series}.

<<CD_lambda_time_series, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=6, fig.cap='Solow residual time series.'>>=
# Add a bogus "none" column that we can use for the no-energy case
m_histData <- historicalData
m_histData$none <- 1.0
# Melt the the historical data to put energy in one column
m_histData <- melt(select(m_histData, Year, iGDP, iK, iL, iXp, iXf, iU, none, Country, Source), 
                   measure.vars = c("none", "iXp", "iXf", "iU"),
                   variable.name = "energyType",
                   value.name = "iE")
# Calculate finite difference approximations of time derivatives for output and all factors of production
fd <- m_histData %>% group_by(Source, Country, energyType) %>% arrange(Year) %>% 
  mutate(dydt = ediff(iGDP, pad="tail") / ediff(Year, pad="tail")) %>%
  mutate(dkdt = ediff(iK, pad="tail") / ediff(Year, pad="tail")) %>%
  mutate(dldt = ediff(iL, pad="tail") / ediff(Year, pad="tail")) %>%
  mutate(dedt = ediff(iE, pad="tail") / ediff(Year, pad="tail"))

origCDKLCoeffs <- subset(AllCoeffs, method=="orig" & model=="cd")
origCDKLCoeffs <- origCDKLCoeffs %>% select(method, country, Source, nestStr, energy, lambda, alpha, beta, gamma)
# Change name of Source REXS to REXS1960 and relevel to conform to information in fdlambda.
origCDKLCoeffs$Source <- as.character(origCDKLCoeffs$Source)
origCDKLCoeffs$Source <- replace(origCDKLCoeffs$Source, which(origCDKLCoeffs$Source=="REXS"), "REXS1960")
origCDKLCoeffs$Source <- as.factor(origCDKLCoeffs$Source)
origCDKLCoeffs$Source <- relevelFactor(origCDKLCoeffs$Source, levs = c("REXS1960", "IST", "Leeds"))

# Fill new columns in the fd data frame with 
# lambda, alpha, beta, and gamma values from the parameter estimation process.
# lambda, alpha, beta, and gamma values for the orig fits are in the origCDKLCoeffs data frame.
# Using the hint at 
# http://stackoverflow.com/questions/15059076/
# r-how-to-call-apply-like-function-on-each-row-of-dataframe-with-multiple-argum
labg <- apply(fd[ , c('Country', 'Source', 'energyType')], 1, function(x){
    row <- as.data.frame(subset(origCDKLCoeffs, 
                                country==x["Country"] & Source==x["Source"] & energy==x["energyType"]))
    abg <- subset(row, select=c(lambda, alpha, beta, gamma))
    if (nrow(abg) < 1){ abg[1, ] <- NA }
    return(abg)
  })
cols <- do.call(rbind, labg)
fd <- cbind(fd, cols)

# Now calculate lambda by finite difference for every year to make a time series
fdlambda <- mutate(fd, lambda_diff = 1/iGDP*dydt - (alpha/iK*dkdt + beta/iL*dldt + gamma/iE*dedt))

# Create a time series graph of the lambda_diff values
ggplot(data=fdlambda, aes(x=Year, y=lambda_diff, color=Source)) + 
  geom_point(size = 1.0) +
  facet_grid(Country ~ energyType) +
  labs(x = "", y = expression(lambda)) + 
#   scale_y_continuous(limits=c(-0.1, 0.1)) +
  xy_theme()
@
